### 2022-12-17 analysis/glc6-lactate.qmd
start work on glc6 fluxes

---
title: "LC-MS ^13^C Lactate Measurements"
date-modified: today
format: html
excecute:
  message: false
---

```{r}
#| label: setup

devtools::load_all()
suppressPackageStartupMessages({
  library(tidyverse)
  library(wmo)
  library(targets)
})

here::i_am("analysis/glc6-lactate.qmd")
here::here()

theme_set(theme_wmo())
```

```{r}
#| label: data

withr::with_dir(here::here(), {
  path <- tar_read(fluxes_glc6_files)
})
```

## Overview

To validate the results of enzymatic lactate assays suggesting decreased glycolysis in hypoxia, cells were treated with [U-^13^C~6~]-glucose (8 mM) in MCDB131 medium supplemented with glutamine (1 mM). DNA samples were collected for PicoGreen assay and medium samples were collected for LC-MS-based analysis of [u-^13^C~3~]-lactate concentration. Medium samples (10 μL) were combined with D~8~-valine (10 μL of 100 100 μM) and extracted with 180 μL MeOH pre-cooled to -80 °C. Samples were spun and the supernatants were analyzed by LC-MS using both full scan and selected ion monitoring with negative polarity. Quality control samples were interspersed across all of the samples to assess for instrument drift.

# Analysis

```{r}
#| label: raw-data
df <-
  read_ms_excel(here::here(path)) |> 
  dplyr::filter(!str_detect(.data$id, "B")) |> 
  tidyr::separate(
    .data$sample, 
    c("type", "number"), 
    sep = "(?<=[[:alpha:]])(?=[[:digit:]])", 
    fill = "left", 
    convert = TRUE
  ) |>
  dplyr::mutate(
    conc = ifelse(is.na(.data$type), .data$number, NA_character_), 
    date = dplyr::case_when(
      .data$type == "A" ~ "2022-11-12", 
      .data$type == "B" ~ "2022-11-17", 
      .data$type == "C" ~ "2022-11-22"
    ), 
    number = ifelse(!is.na(conc), NA, .data$number), 
    batch = "b", 
    cell_type = "lf", 
    experiment = "substrate", 
    type = dplyr::case_when(
      .data$type == "QC" ~ "qc", 
      is.na(.data$type) ~ "std", 
      TRUE ~ "sample"
    ), 
    detector = "lcms", 
    run = "a"
  ) |> 
  dplyr::select(
    "metabolite", 
    "detector", 
    "mz", 
    "rt", 
    "cell_type",
    "experiment", 
    "batch", 
    "date", 
    "run", 
    "id", 
    "type", 
    "number", 
    "conc", 
    "area"
  )
```

## Drift

```{r}
#| label: drift-plot

df_drift <- 
  df |> 
  filter(type == "qc") |> 
  group_by(metabolite) |> 
  mutate(norm = area / mean(area))

ggplot(df_drift) +
  aes(
    x = number, 
    y = norm, 
    color = metabolite
  ) +
  geom_line() + 
  labs(
    x = "Run order", 
    y = "Normalized peak area", 
    color = "Metabolite", 
    title = "QC Drift"
  )
```

The QC peak areas suggest that correcting for experimental drift may be helpful over the time course of the experiment.

```{r}
#| label: qc-variance

ggplot(df_drift) +
  aes(
    x = metabolite, 
    y = norm, 
    color = metabolite
  ) +
  geom_boxplot(show.legend = FALSE) +
  geom_jitter(
    width = 0.1, 
    show.legend = FALSE
  ) +
  labs(
    x = "Metabolite", 
    y = "Normalized peak area", 
    title = "QC Variance"
  )
```

Regardless of drift, the QC peak areas are quite similar across the experimental time course.

```{r}
#| label: se

se <- 
  df |> 
  new_tbl_se(
    a_data = "area", 
    f_names = "metabolite", 
    f_data = c("metabolite", "detector", "mz", "rt"), 
    s_names = "id", 
    s_data = c(
      "cell_type", 
      "experiment", 
      "batch", 
      "run", 
      "date", 
      "type", 
      "number", 
      "conc"
    )
  ) |> 
  tbl_to_se()

se_drift <- 
  correct_drift(se) |> 
  se_to_tbl(rownames = "metabolite")
```

```{r}
#| label: drift-corr-sanity-check

se_drift |> 
  filter(type == "qc") |> 
  group_by(metabolite) |> 
  mutate(norm = area / mean(area)) |> 
  ggplot() +
  aes(
    x = number, 
    y = norm, 
    color = metabolite
  ) +
  geom_line() + 
  labs(
    x = "Run order", 
    y = "Normalized peak area", 
    color = "Metabolite", 
    title = "Drift Corrected"
  )
```

```{r}
#| label: qbias-corr

se_drift |> 
  dplyr::filter(type != "qc")
```

## Standard Curves

```{r}
#| label: standard-curves

se_drift |> 
  dplyr::filter(type == "std" & metabolite == "LAC M0") |> 
  dplyr::mutate(conc = as.numeric(conc)) |> 
  ggplot() +
  aes(
    x = conc, 
    y = area
  ) +
  geom_point()
```

## Samples

```{r}
#| label: samples

pheno <- 
  tibble::tibble(
    number = 1:24, 
    oxygen = rep(c("21%", "0.5%"), times = 4, each = 3), 
    time = rep(c(0, 24, 48, 72), each = 6), 
    rep = rep(c("x", "y", "z"), 8)
  )

se_drift |> 
  filter(type == "sample") |> 
  filter(metabolite == "LAC M3") |> 
  left_join(pheno, by = "number") |> 
  group_by(date, oxygen, time) |> 
  wmo::remove_nested_outliers(area, remove = TRUE) |> 
  summarize(area = mean(area)) |> 
  ggplot() +
  aes(
    x = time, 
    y = area, 
    color = oxygen
  ) +
  geom_point() +
  stat_summary(
    geom = "line"
  )
```

These plots suggest that, for the 100 μL extraction buffer, the maximum cell count is ~ 250,000. We will exclude the 300,000 value prior to calculating the slopes of the curves. For the PASMC samples, the 400,000 cell sample is quite high compared to the other samples. I will also remove these values prior to calculation of the slopes.