### 2022-12-21 analysis/glc6-lactate.qmd
fix integration of GLC6 data into final flux measurements

---
title: "LC-MS ^13^C Lactate Measurements"
date-modified: today
format: html
excecute:
  message: false
---

```{r}
#| label: setup
#| output: false

devtools::load_all()
suppressPackageStartupMessages({
  library(tidyverse)
  library(wmo)
  library(targets)
})

here::i_am("analysis/glc6-lactate.qmd")
here::here()

theme_set(theme_wmo())
```

```{r}
#| label: data

withr::with_dir(here::here(), {
  se_raw <- tar_read(fluxes_glc6_se)
  df_drift <- tar_read(fluxes_glc6_drift)
  std <- tar_read(conc_std)
  fluxes <- tar_read(fluxes)
})
```

## Overview

To validate the results of enzymatic lactate assays suggesting decreased glycolysis in hypoxia, cells were treated with [U-^13^C~6~]-glucose (8 mM) in MCDB131 medium supplemented with glutamine (1 mM). DNA samples were collected for PicoGreen assay and medium samples were collected for LC-MS-based analysis of [u-^13^C~3~]-lactate concentration. Medium samples (10 μL) were combined with D~8~-valine (10 μL of 100 100 μM) and extracted with 180 μL MeOH pre-cooled to -80 °C. Samples were spun and the supernatants were analyzed by LC-MS using both full scan and selected ion monitoring with negative polarity. Quality control samples were interspersed across all of the samples to assess for instrument drift.

# Analysis

## Drift Correction

```{r}
#| label: drift-plot

df_raw <- 
  se_raw |> 
  se_to_tbl(rownames = "metabolite") |> 
  filter(type == "qc") |> 
  group_by(metabolite) |> 
  mutate(norm = area / mean(area))

ggplot(df_raw) +
  aes(
    x = number, 
    y = norm, 
    color = metabolite
  ) +
  geom_line() + 
  labs(
    x = "Run order", 
    y = "Normalized peak area", 
    color = "Metabolite", 
    title = "QC Drift"
  )
```

The QC peak areas suggest that correcting for experimental drift may be helpful over the time course of the experiment.

```{r}
#| label: qc-variance

ggplot(df_raw) +
  aes(
    x = metabolite, 
    y = norm, 
    color = metabolite
  ) +
  geom_boxplot(show.legend = FALSE) +
  geom_jitter(
    width = 0.1, 
    show.legend = FALSE
  ) +
  labs(
    x = "Metabolite", 
    y = "Normalized peak area", 
    title = "QC Variance"
  )
```

Regardless of drift, the QC peak areas are quite similar across the experimental time course.

```{r}
#| label: drift-corr-sanity-check

df_drift |> 
  filter(type == "qc") |> 
  group_by(metabolite) |> 
  mutate(norm = area / mean(area)) |> 
  ggplot() +
  aes(
    x = number, 
    y = norm, 
    color = metabolite
  ) +
  geom_line() + 
  labs(
    x = "Run order", 
    y = "Normalized peak area", 
    color = "Metabolite", 
    title = "Drift Corrected"
  )
```

## Standard Curves

```{r}
#| label: standard-curves
#| results: hide

std |> 
  filter(experiment == "substrate" & metabolite == "lactate") |> 
  pull(plots)
```

## Samples

```{r}
#| label: samples

fluxes |> 
  filter(experiment == "substrate")
df_drift |> 
  filter(type %in% letters[1:3]) |> 
  filter(metabolite == "LAC M0") |> 
  left_join(pheno, by = "number") |> 
  group_by(type, oxygen, time) |> 
  wmo::remove_nested_outliers(area, remove = TRUE) |> 
  summarize(area = mean(area)) |> 
  ggplot() +
  aes(
    x = time, 
    y = area, 
    color = oxygen
  ) +
  geom_point() +
  stat_summary(
    geom = "line"
  )
```

These plots suggest that, for the 100 μL extraction buffer, the maximum cell count is ~ 250,000. We will exclude the 300,000 value prior to calculating the slopes of the curves. For the PASMC samples, the 400,000 cell sample is quite high compared to the other samples. I will also remove these values prior to calculation of the slopes.