### 2023-02-10 analysis/flux-differences.qmd
add model flux analysis

---
title: "Flux Differences"
date-modified: today
format: html
excecute:
  message: false
---

```{r}
#| label: setup
#| output: false

devtools::load_all()
suppressPackageStartupMessages({
  library(tidyverse)
  library(wmo)
  library(targets)
})

here::i_am("analysis/flux-differences.qmd")

theme_set(theme_wmo())
```

```{r}
#| label: data

withr::with_dir(here::here(), {
  model_fluxes <- tar_read(model_fluxes)
  flux_differences <- tar_read(map_flux_differences)
})
```

# Overview

In this document, we will analyze the output from the metabolic flux maps.

# Analysis

## Significant Differences

Identify reactions that are significantly different based on whether the confidence intervals overlap. This is more conservative than p < 0.05.

```{r}
#| label: functions

calculate_ratio <- function(norm) {
  flux_differences |> 
    filter(normalization == norm) |> 
    # filter(!is.na(ratio)) |> 
    select(-c(normalization, equation, ctl, contains(c("flux", "lb", "ub")))) |> 
    pivot_wider(names_from = exp, values_from = ratio)
}

print_flux_table <- function(df, caption) {
  df |> 
    # mutate(across(c(`0.5%`, BAY), ~as.character(signif(., 3)))) |> 
    gt::gt() |> 
    gt::cols_label(
      cell_type = "Cell type", 
      pathway = "Pathway", 
      index = "Index", 
      id = "ID", 
      type = "Type", 
      `0.5%` = "0.5%", 
    ) |> 
    gt::fmt_scientific(c(`0.5%`, BAY)) |> 
    gt::tab_header(title = caption) |> 
    gt::cols_align(
      "left", 
      c(pathway, id, type)
    ) |> 
    gt::opt_all_caps() |> 
    gt::opt_align_table_header(align = "left") |> 
    gt::tab_options(row_group.as_column = TRUE)
}
```

```{r}
#| label: significant-flux-differences

calculate_ratio("none") |> 
  print_flux_table("Significant Flux Differences")
```

Do any of these differences have a negative ratio?

```{r}
#| label: reversed-fluxes

calculate_ratio("none") |> 
  filter(`0.5%` < 0 | BAY < 0) |> 
  print_flux_table("Reversed fluxes")
```

For both of these models, some of the PPP fluxes reverse.

## Normalized differences

It doesn't look like there is any dramatic reprogramming of the intracellular metabolic pathways secondary to hypoxia or BAY treatment. In other words, most of the differences are related to absolute rates rather than differential usage. Another way to look at this might be to normalize the fluxes to glucose uptake or biomass fluxes. This would show all fluxes in terms of total glucose use or growth rate, for example.

```{r}
#| label: growth-normalized

calculate_ratio("growth") |> 
  print_flux_table("Significant differences in growth-normalized fluxes")
```

```{r}
#| label: glucose-normalized

calculate_ratio("glucose") |> 
  print_flux_table("Significant differences in glucose-normalized fluxes")
```